\subsection{\bap Typing}
\label{vinesec:typecheck}
\input{typecheck-table}

The \bap IL syntax described in Table~\ref{vine:syntax} is designed
to be simple. A literal interpretation, however, allows several
nonsensical operations such as left-shifting two memory variables.  We
type-check \bap programs using the  rules shown in
Table~\ref{vine:typecheck} to weed out obviously nonsensical
operations.  We emphasize that type-checking to show program safety is outside
the scope of \bap.

The type-checking rules are straight-forward.  \bap requires that
binary and unary operations be on scalars of the same register type
(except shifts where the shift amount need only be an integer). We
type-check {\tt hi} and {\tt low} casts such that the resulting type
is a smaller register type (in terms of the sub-typing relationship
{\sc reg-subtype}), as expected.  {\tt unsigned} and {\tt signed}
casts must be to a wider register type. One issue with un-normalized
memory is that it may have multi-byte accesses, e.g., a store on x86
could be of 32, 16, or 8 bits. Normalized memory, on the other hand,
only allows stores and loads of bytes.  We define an auxiliary
predicate ``mem\_compat'' to ensure that normalized memory is accessed
appropriately.  Also note that as a convenience we assume 64-bit
addressing in the rules, but allow smaller addressable memories via
the {\tt reg-subtype} sub-typing rule.

A program type-checks if all instructions type-check under the
declared variable types.  We require all variables be declared because
explicitly typed languages are easier to check. 

